Binarity's Biases on Transit Survey Occurrence Rates

### 2018-02-24 paper/main.tex
aesthetic improvements

\title{ Binarity's Biases on Transit Survey Occurrence Rates }
%
\correspondingauthor{L. Bouma}
\email{luke@astro.princeton.edu}
%
\author[0000-0002-0514-5538]{L. G. Bouma}
\affiliation{
    Department of Astrophysical Sciences,
    Princeton University,
    4 Ivy Lane, Princeton, NJ 08540, USA}
\author[0000-0003-1298-9699]{K. Masuda}
\affiliation{
    Department of Astrophysical Sciences,
    Princeton University,
    4 Ivy Lane, Princeton, NJ 08540, USA}
\author[0000-0002-4265-047X]{J. N. Winn}
\affiliation{
    Department of Astrophysical Sciences,
    Princeton University,
    4 Ivy Lane, Princeton, NJ 08540, USA}
%
%
\begin{abstract}
%
Wide-field surveys for transiting planets, such as the NASA {\it
  Kepler} and {\it TESS} missions, are usually conducted without
knowing which stars have binary companions. Unresolved and
unrecognized binaries give rise to systematic errors in planet
occurrence rates, including misclassified planets and mistakes in
completeness corrections.  The individual errors can have different
signs, making it difficult to anticipate the net effect on inferred
occurrence rates. Here we use simplified models of signal-to-noise
limited transit surveys to try and clarify the situation. We derive a
formula for the apparent occurrence rate density measured by an
observer who falsely assumes all stars are single. The formula depends
on the binary fraction; the mass function of the secondary stars; and
the true occurrence of planets around primaries, secondaries, and
single stars. It also takes into account the Malmquist bias by which
binaries are over-represented in flux-limited samples. Application of
the formula to an idealized {\it Kepler}-like survey shows that for
planets larger than 2\,$R_\oplus$, the net systematic error is of
order 5\%.  In particular, unrecognized binaries are unlikely to be
the reason for the apparent discrepancies between hot Jupiter
occurrence rates measured in different surveys.  For smaller planets
the size the errors are potentially larger: the occurrence of
Earth-sized planets could be overestimated by as much as 50\%. We also
show that whenever high-resolution imaging reveals a transit host star
to be a binary, the planet is usually more likely to orbit the primary
star than the secondary star.
%
\end{abstract}
%
\keywords{
    methods: data analysis ---
    planets and satellites: detection ---
    surveys}
%
%

@@

\begin{figure}[!tb]
    \begin{center}
        \includegraphics[width=.6\textwidth]{figures/occ_rate_vs_radius_model_1_brokenx.pdf}
    \end{center}
    \vspace{-0.5cm}
    \caption{ Apparent occurrence rate $\Lambda\a$, and true
    occurrence rate $\Lambda$ in a universe with only one type of
    star, one type of planet (with radius $R_\oplus$), and a number
    density of binary stars 5\% the number density of single stars.
    The occurrence
    rate of planets with radius $R_\oplus$ is underestimated, and the
    occurrence rate of planets with radius
    $R_\oplus/\sqrt{2}$ is overestimated.}
    \label{fig:occ_rate_model_1}
\end{figure}

@@

\begin{figure}[!t]
    \centering
    \epsscale{1.15}
    \plottwo{figures/int_rate_density_vs_radius_model_3_rpu_22.5_manyZs.pdf}{figures/int_occ_rate_vs_radius_model_3_rpu_22.5_manyZs.pdf}
    \caption{ {\it Left:} Apparent occurrence rate density
      ($\Gamma\a$), compared to the true occurrence rate density
      ($\Gamma_0$) for single stars and primary stars in binaries.
      Three different cases are plotted, with different choices for
      the true occurrence rate density for secondary stars relative to
      that for single stars ($Z_2/Z_0$).
      {\it Right:} Same, but for apparent occurrence
      rate ($\Lambda\a$) within $0.5\,R_\oplus$ bins.
      In both cases the true planet
      radius distribution is specified by
      Equation~\ref{eq:model3_radius_distribution}.
    }
    \label{fig:occ_rate_model_3_log}
\end{figure}

@@

\begin{figure}[!t]
    \centering
    \epsscale{1.15}
    \plottwo{figures/int_rate_density_vs_radius_model_4_xcut_manyZs.pdf}{figures/int_occ_rate_vs_radius_model_4_xcut_manyZs.pdf}
    \caption{ {\it Left:} True and apparent occurrence rate densities,
      and {\it right:} occurrence rates, as a function of planet
      radius, using a model for the radius distribution that exhibits
      a gap between 1.5 and 2\,$R_\oplus$
      (Equation~\ref{eq:model4_radius_distribution}). }
    \label{fig:model_4}
\end{figure}

@@

\paragraph{Number of detected planets}
Inserting the results into Equation~\ref{eq:n_det}, we find
\begin{align}
    \notag
    n_{\rm det}^0(r\a, M\a)
    &=\int\mathrm{d}r\,\mathrm{d}M~\mathcal{N}_\star^0(r\a, M\a; r, M)
    \, \Gamma_0(r, M) \, p_{\rm tra}(M)\\
    &=N_\star^0(r\a, M\a)\, \Gamma_0(r\a, M\a) \, p_{\rm 
        tra}(M\a),\\
    \notag
    n_{\rm det}^1(r\a, M\a)
    &=\int\mathrm{d}r\,\mathrm{d}M~\mathcal{N}_\star^1(r\a, M\a; r, M)
    \, \Gamma_1(r, M) \, p_{\rm tra}(M)\\
    &=N_\star^0(r\a, M\a)\, p_{\rm tra}(M\a) \,
    \frac{n\b}{n\s}
    \int \mathrm{d}q \,
         \mathcal{D}_1^3 f(q)\cdot
         \mathcal{D}_1 \Gamma_1(\mathcal{D}_1r\a, M\a),\\
    n_{\rm det}^2(r\a, M\a)
    \notag
    &=\int\mathrm{d}r\,\mathrm{d}M~\mathcal{N}_\star^2(r\a, M\a; r, M)
    \, \Gamma_2(r, M) \, p_{\rm tra}(M)\\
    &=
    N_\star^0(r\a, M\a)\, p_{\rm tra}(M\a) \, \frac{n\b}{n\s}
    \int {\rm d}q\, 
         \mathcal{D}_1^3 f(q)\cdot q \mathcal{D}_2
         \Gamma_2(\mathcal{D}_2r\a, qM\a)\cdot
         \frac{R(qM\a)}{R(M\a)} q^{-1/3}.
    \label{eq:n_det_2}
\end{align}

>>>>

\title{ Binarity's Biases on Transit Survey Occurrence Rates }
%
\correspondingauthor{L. Bouma}
\email{luke@astro.princeton.edu}
%
\author[0000-0002-0514-5538]{L. G. Bouma}
\affiliation{
    Department of Astrophysical Sciences,
    Princeton University,
    4 Ivy Lane, Princeton, NJ 08540, USA}
\author[0000-0003-1298-9699]{K. Masuda}
\affiliation{
    Department of Astrophysical Sciences,
    Princeton University,
    4 Ivy Lane, Princeton, NJ 08540, USA}
\author[0000-0002-4265-047X]{J. N. Winn}
\affiliation{
    Department of Astrophysical Sciences,
    Princeton University,
    4 Ivy Lane, Princeton, NJ 08540, USA}
%
%
\begin{abstract}
%
Wide-field surveys for transiting planets, such as the NASA {\it
  Kepler} and {\it TESS} missions, are usually conducted without
knowing which stars have binary companions. Unresolved and
unrecognized binaries give rise to systematic errors in planet
occurrence rates, including misclassified planets and mistakes in
completeness corrections.  The individual errors can have different
signs, making it difficult to anticipate the net effect on inferred
occurrence rates. Here we use simplified models of signal-to-noise
limited transit surveys to try and clarify the situation. We derive a
formula for the apparent occurrence rate density measured by an
observer who falsely assumes all stars are single. The formula depends
on the binary fraction; the mass function of the secondary stars; and
the true occurrence of planets around primaries, secondaries, and
single stars. It also takes into account the Malmquist bias by which
binaries are over-represented in flux-limited samples. Application of
the formula to an idealized {\it Kepler}-like survey shows that for
planets larger than 2\,$R_\oplus$, the net systematic error is of
order 5\%.  In particular, unrecognized binaries are unlikely to be
the reason for the apparent discrepancies between hot Jupiter
occurrence rates measured in different surveys.  For smaller planets
the errors are potentially larger: the occurrence of Earth-sized
planets could be overestimated by as much as 50\%. We also show that
whenever high-resolution imaging reveals a transit host star to be a
binary, the planet is usually more likely to orbit the primary star
than the secondary star.
%
\end{abstract}
%
\keywords{
    methods: data analysis ---
    planets and satellites: detection ---
    surveys}
%
%

@@

\begin{figure}[!tb]
    \begin{center}
        \includegraphics[width=.6\textwidth]{figures/occ_rate_vs_radius_model_1_brokenx.pdf}
    \end{center}
    \vspace{-0.5cm}
    \caption{ Apparent occurrence rate $\Lambda\a$, and true
    occurrence rate $\Lambda$ in a universe with only one type of
    star, one type of planet (with radius $R_\oplus$), and 
    a binary system number density 5\% that of single stars.
    The occurrence
    rate of planets with radius $R_\oplus$ is underestimated, and the
    occurrence rate of planets with radius
    $R_\oplus/\sqrt{2}$ is overestimated.}
    \label{fig:occ_rate_model_1}
\end{figure}

@@

\begin{figure}[!t]
    \centering
    \epsscale{1.15}
    \plottwo{figures/int_rate_density_vs_radius_model_3_rpu_22.5_manyZs.pdf}{figures/int_occ_rate_vs_radius_model_3_rpu_22.5_manyZs.pdf}
    \caption{ {\it Left.}---Apparent occurrence rate density
      ($\Gamma\a$), compared to the true occurrence rate density
      ($\Gamma_0$) for single stars and primary stars in binaries.
      Three different cases are plotted, with different choices for
      the number of planets per secondary star relative to
      that for single stars ($Z_2/Z_0$).
      {\it Right.}---Same, but for apparent occurrence
      rate ($\Lambda\a$) within $0.5\,R_\oplus$ bins.
      In both cases the true planet
      radius distribution is specified by
      Equation~\ref{eq:model3_radius_distribution}.
    }
    \label{fig:occ_rate_model_3_log}
\end{figure}

@@

\begin{figure}[!t]
    \centering
    \epsscale{1.15}
    \plottwo{figures/int_rate_density_vs_radius_model_4_xcut_manyZs.pdf}{figures/int_occ_rate_vs_radius_model_4_xcut_manyZs.pdf}
    \caption{ {\it Left.}---True and apparent occurrence rate densities,
      and {\it right.}---occurrence rates, as a function of planet
      radius, using a model for the radius distribution that exhibits
      a gap between 1.5 and 2\,$R_\oplus$
      (Equation~\ref{eq:model4_radius_distribution}). }
    \label{fig:model_4}
\end{figure}

@@

\paragraph{Number of detected planets}
Inserting the results into Equation~\ref{eq:n_det}, we find
\begin{align}
    \notag
    n_{\rm det}^0(r\a, M\a)
    &=\int\mathrm{d}r\,\mathrm{d}M~\mathcal{N}_\star^0(r\a, M\a; r, M)
    \, \Gamma_0(r, M) \, p_{\rm tra}(M)\\
    &=N_\star^0(r\a, M\a)\, \Gamma_0(r\a, M\a) \, p_{\rm 
        tra}(M\a),
    \label{eq:n_det_0}
    \\
    \notag
    n_{\rm det}^1(r\a, M\a)
    &=\int\mathrm{d}r\,\mathrm{d}M~\mathcal{N}_\star^1(r\a, M\a; r, M)
    \, \Gamma_1(r, M) \, p_{\rm tra}(M)\\
    &=N_\star^0(r\a, M\a)\, p_{\rm tra}(M\a) \,
    \frac{n\b}{n\s}
    \int \mathrm{d}q \,
         \mathcal{D}_1^3 f(q)\cdot
         \mathcal{D}_1 \Gamma_1(\mathcal{D}_1r\a, M\a),\\
    n_{\rm det}^2(r\a, M\a)
    \notag
    &=\int\mathrm{d}r\,\mathrm{d}M~\mathcal{N}_\star^2(r\a, M\a; r, M)
    \, \Gamma_2(r, M) \, p_{\rm tra}(M)\\
    &=
    N_\star^0(r\a, M\a)\, p_{\rm tra}(M\a) \, \frac{n\b}{n\s}
    \int {\rm d}q\, 
         \mathcal{D}_1^3 f(q)\cdot q \mathcal{D}_2
         \Gamma_2(\mathcal{D}_2r\a, qM\a)\cdot
         \frac{R(qM\a)}{R(M\a)} q^{-1/3}.
    \label{eq:n_det_2}
\end{align}
